\section{7 January 2014}
\subsection{Stable Marriage Problem}
Gale-Shapley, 1962\\

Residents list hospital preferences, hospitals list resident preferences.\\

Try to find a \emph{stable} arrangement. That is, there does not exist a resident $R$ or hospital $H$ such that $R$ and $H$ are not currently assigned to each other, but would prefer to be assigned to each other over their current assignments.

\begin{figure}[h]
\begin{minipage}[h]{.5\linewidth}
\centering
\begin{tabular}{c|c c c}
  & 1 & 2 & 3 \\ \hline
A & Y & X & Z \\
B & X & Y & Z \\
C & Z & X & Y
\end{tabular}
\subcaption{Amy, Bertha, Clare}
\end{minipage}
\begin{minipage}[h]{.5\linewidth}
\centering
\begin{tabular}{c|c c c}
  & 1 & 2 & 3 \\ \hline
X & A & B & C \\
Y & B & C & A \\
Z & B & A & C
\end{tabular}
\subcaption{Xavier, Yancey, Zeus}
\end{minipage}
\end{figure}

Do stable matchings exist? If so, how do we find them?\\

(X, A) (Y, B): stable\\
(X, A) (Z, C): stable\\
(Y, B) (Z, C): stable\\

(X, C), (Y, B), and (Z, C) is \emph{unstable}: B prefers X over Y, X prefers B over C.

\subsubsection*{Possible approaches}
\begin{description}
\item[Approach 1] Assign a value to each pair and pick match with largest value
\item[Approach 2] If a man and woman prefer each other, match them and repeat
\item[Approach 3] If exists unique top choice, match them. If two or more have same top choice, go through the others and identify stable pairs.
\end{description}

\subsubsection*{Gale{\textendash}Shapley Method}
\begin{algorithmic}
\While{there exists unmatched man $m$ who has not proposed to all women}
\State{$m$ proposes to next woman $w$ on his list}
\If{$w$ is unmatched or $w$ is engaged to $m'$ but prefers $m$ to $m'$}
\State{$w$ gets matched to $m$ (dumping $m'$ if necessary)}
\Else{$w$ rejects $m$}
\EndIf
\EndWhile
\end{algorithmic}

\subsubsection*{Example}
\begin{multicols}{2}
X proposes to A\\
Z proposes to B\\
Y proposes to B\\
Z proposes to A, A rejects Z\\
Z proposes to C
\columnbreak

(X, A)\\
(X, A), (Z, B)\\
(X, A), \sout{(Z, B)} (Y, B)\\
(X, A), (Y, B)\\
(X, A), (Y, B), (Z, C)
\end{multicols}

\begin{description}
\item[Lemma 1]: G{\textendash}S always terminates.
\item[Lemma 2]: At termination, the matching is stable.
\end{description}

\subsubsection*{Stable roommates}
\begin{tabular}{c|c c c}
  & 1 & 2 & 3\\ \hline
A & B & C & D\\
B & C & A & D\\
C & A & B & D\\
D & A & B & C
\end{tabular}

Never stable, because no one wants to live with D.\\

\emph{Observation 1:} Once a woman is proposed to, she is always matched.\\
\emph{Proof:} First time a woman is proposed to (line 2) she gets engaged (line 4). Thereafter she breaks off the engagement only to get engaged to someone else.\\

\emph{Observation 2:} At termination, every man is either matched or has proposed to every woman.\\
\emph{Proof of Lemma 1:}
\begin{itemize}
\item In each iteration of the while loop, some man $m$ proposes to some woman $w$.
\item No man proposes to the same woman twice.
\item No more than $n^2$ proposals.
\item No more than $n^2$ iterations.
\item G{\textendash}S terminates in $\leq n^2$ iterations.
\end{itemize}